Stabilization-Free Virtual Element Methods

### 2023-03-10 stabfreevem_sinum.tex

The key to construct stabilization-free VEMs is to find a finite-dimensional space $\mathbb{V}(K)$ for polytope $K$ and a projector $Q_K$ onto the space $\mathbb{V}(K)$ such that
\begin{enumerate}[(C1)]
\item It holds the norm equivalence 
\begin{equation}\label{intro:gradVknormequiv} 
\|Q_{K}\nabla v\|_{0,K}\eqsim \|\nabla v\|_{0,K} \quad \forall~v\in V_k(K)
\end{equation}
on shape function space $V_k(K)$ of virtual elements;
\item The projection $Q_{K}\nabla v$ is computable based on the degrees of freedom (DoFs) of virtual elements for $v\in V_k(K)$.
\end{enumerate}
\XH{The hidden constants in \eqref{intro:gradVknormequiv} are independent of the size of $K$, but depend on the degree of polynomials, and the chunkiness parameter and the geometric dimension of $K$; see Section~\ref{sec:meshcondition} for details.}
We can choose $Q_{K}$ as the $L^2$-orthogonal projector with respect to the
inner product $(\cdot, \cdot)_K$. The norm equivalence
\eqref{intro:gradVknormequiv} implies that the space $\mathbb{V}(K)$ should be
sufficiently large compared with the virtual element space $V_k(K)$.  In
standard virtual element methods, $Q_{k-1}^{K}\nabla v$
\cite{BeiraodaVeigaBrezziMariniRusso2016} or $\nabla\Pi_k^{K}v$
\cite{BeiraoBrezziCangianiManziniEtAl2013,BeiraoBrezziMariniRusso2014,AhmadAlsaediBrezziMariniEtAl2013,AyusodeDiosLipnikovManzini2016}
are used, where $Q_{k-1}^{K}$ is the $L^2$-orthogonal projector onto the
$(k-1)$-th order polynomial space $\mathbb P_{k-1}(K; \mathbb{R}^d)$, and
$\Pi_k^{K}$ is the $H^1$ projection operator onto the $k$-th order polynomial
space $\mathbb P_{k}(K)$.  While only
\[
\|Q_{k-1}^{K}\nabla v\|_{0,K}\lesssim \|\nabla v\|_{0,K}, \quad \|\nabla\Pi_k^{K}v\|_{0,K}\lesssim \|\nabla v\|_{0,K}
\]
hold
rather than the norm equivalence \eqref{intro:gradVknormequiv}, then the additional stabilization term is usually necessary to ensure the coercivity of the discrete bilinear form.
To remove the additional stabilization term, $\mathbb{V}(K)$ is taken as $\mathbb P_{l}(K; \mathbb{R}^d)$ with large enough $l$ even for the lowest order case $k=1$ in \cite{BerroneBorioMarcon2021,BerroneBorioMarcon2022,DAltriMirandaPatrunoSacco2021}, and the virtual element space $V_k(K)$ has to be modified accordingly to make $Q_{l}^{K}\nabla v$ computable.
Instead, \XH{based on a regular simplicial tessellation of polytope $K$}, we employ $k$-th order or $(k-1)$-th order $H(\div)$-conforming macro finite elements as $\mathbb{V}(K)$ in this paper, and keep the virtual element space $V_k(K)$ as the usual ones.

@@

\subsection{Notation}
Let $\Omega\subset \mathbb{R}^d$ be a bounded
polytope. Given a bounded domain $K\subset\mathbb{R}^{d}$ and a
non-negative integer $m$, let $H^m(K)$ be the usual Sobolev space of functions
on $K$.
The corresponding norm and semi-norm are denoted respectively by
$\Vert\cdot\Vert_{m,K}$ and $|\cdot|_{m,K}$. By convention, let $L^2(K)=H^0(K)$. Denote by $L_0^2(K)$ the space of all functions in $L^2(K)$ with zero integral.
Let $(\cdot, \cdot)_K$ be the standard inner product on $L^2(K)$. If $K$ is $\Omega$, we abbreviate
$\Vert\cdot\Vert_{m,K}$, $|\cdot|_{m,K}$ and $(\cdot, \cdot)_K$ by $\Vert\cdot\Vert_{m}$, $|\cdot|_{m}$ and $(\cdot, \cdot)$,
respectively. Let $H_0^m(K)$ be the closure of $\mathcal C_{0}^{\infty}(K)$ with
respect to the norm $\Vert\cdot\Vert_{m,K}$, and $L_0^2(K)$ consist of all functions in $L^2(K)$ with zero mean value.
For integer $k\geq0$,
notation $\mathbb P_k(K)$ stands for the set of all
polynomials over $K$ with the total degree no more than $k$. Set $\mathbb P_{-1}(K)=\{0\}$.
For a Banach space $B(K)$, let $B(K; \mathbb{X}):=B(K)\otimes\mathbb{X}$ with $\mathbb{X}=\mathbb{R}^d$ and $\mathbb{K}$ being the set of antisymmetric matrices.
Denote by $Q_k^{K}$ the $L^2$-orthogonal projector onto $\mathbb P_k(K)$ or $\mathbb P_{k}(K; \mathbb{X})$.
For tensor $\boldsymbol{\tau}$, let $\skw\boldsymbol{\tau}:=(\boldsymbol{\tau}-\boldsymbol{\tau}^{\intercal})/2$ be the antisymmetric part of $\boldsymbol{\tau}$.
Denote by $\#S$ the number of elements in a finite set $S$.

@@

By (3.5) in \cite{ChenHuang2021divdiv}, we have the direct decomposition
\begin{equation}\label{eq:Pkvectordecomp}  
\mathbb P_{k-1}(T;\mathbb R^{d})=\nabla\mathbb P_k(T)\oplus\mathbb P_{k-2}(T;\mathbb K)\boldsymbol{x},
\end{equation}
\XH{where $\mathbb P_{k-2}(T;\mathbb K)\boldsymbol{x}:=\{\boldsymbol{\tau}\boldsymbol{x}: \boldsymbol{\tau}\in \mathbb P_{k-2}(T;\mathbb K)\}$ with $\boldsymbol{x}\in T$ being the independent variable.}
\XH{Recall (35) in \cite{Chen;Huang:2020Finite} that
\begin{equation}\label{eq:20230205}
\mathbb P_{k}(T)\cap\ker(I+\boldsymbol{x}\cdot\nabla)=\{0\}.
\end{equation}
Indeed, \eqref{eq:20230205} follows from the Euler's formula \cite[(2.1)]{ChenHuang2021divdiv}
$$
\boldsymbol{x}\cdot\nabla q=kq\quad\forall~q\in \mathbb H_k(T),
$$
where $\mathbb H_k(T)$ is the space of homogeneous polynomials of degree $k$.}

@@

The rectangular domain $\Omega$ is partitioned by the convex polygon mesh
$\mathcal T_0$ and non-convex polygon mesh $\mathcal T_1$, respectively, 
In both virtual element methods \eqref{eq:vem} and \eqref{eq:cfmvem}, we choose $k = 1, 2, 5$.
The numerical results of the nonconforming virtual element method \eqref{eq:vem} on meshes $\mathcal T_0$ and $\mathcal T_1$ mesh are shown in Figure \ref{fig:rate1}. We can see that $\|u - Q_h u_h\|_0=O(h^{k+1})$ and $\|\nabla u - Q_{h, k-1}^{\div}\nabla_h u_h\|_0=O(h^{k})$, which coincide with Theorem~\ref{thm:errorestimateH1}.
And the numerical results of the conforming virtual element method \eqref{eq:cfmvem} are presented in Figure \ref{fig:rate2}. Again $\|u - Q_h u_h\|_0=O(h^{k+1})$ and $\|\nabla u - Q_{h, k}^{\div}\nabla u_h\|_0=O(h^{k})$, which confirm the
theoretical convergence rate in Theorem~\ref{thm:cfmerrorestimateH1}.
\begin{figure}[htbp]
\centering
\begin{minipage}[t]{0.49\linewidth}
\centering
\includegraphics[width=6cm]{./figures/ncvem_convex.pdf}
%\caption{Numberical result of example with $\mathcal T_0$.}
\end{minipage}%
\begin{minipage}[t]{0.49\linewidth}
\centering
\includegraphics[width=6cm]{./figures/ncvem_nonconvex.pdf}
%\caption{[Numberical result of example with $\mathcal T_1$].}
\end{minipage}%
\centering
\caption{Errors $\|u - Q_h u_h\|_0$ and $\|\nabla u - Q_{h, k-1}^{\div}\nabla_h u_h\|_0$ 
of nonconforming virtual element method \eqref{eq:vem} on 
$\mathcal T_0$(left) and $\mathcal T_1$(right) with $k=1, 2, 5$.}
\label{fig:rate1}
\end{figure}

>>>>

The key to construct stabilization-free VEMs is to find a finite-dimensional space $\mathbb{V}(K)$ for polytope $K$ and a projector $Q_K$ onto the space $\mathbb{V}(K)$ such that
\begin{enumerate}[(C1)]
\item It holds the norm equivalence 
\begin{equation}\label{intro:gradVknormequiv} 
\|Q_{K}\nabla v\|_{0,K}\eqsim \|\nabla v\|_{0,K} \quad \forall~v\in V_k(K)
\end{equation}
on shape function space $V_k(K)$ of virtual elements;
\item The projection $Q_{K}\nabla v$ is computable based on the degrees of freedom (DoFs) of virtual elements for $v\in V_k(K)$.
\end{enumerate}
\XH{The hidden constants in \eqref{intro:gradVknormequiv} are independent of the size of $K$, but depend on the degree of polynomials, and the chunkiness parameter and the geometric dimension of $K$; see Section~\ref{sec:meshcondition} for details.}
We can choose $Q_{K}$ as the $L^2$-orthogonal projector with respect to the
inner product $(\cdot, \cdot)_K$. The norm equivalence
\eqref{intro:gradVknormequiv} implies that the space $\mathbb{V}(K)$ should be
sufficiently large compared with the virtual element space $V_k(K)$.  In
standard virtual element methods, $Q_{k-1}^{K}\nabla v$
\cite{BeiraodaVeigaBrezziMariniRusso2016} or $\nabla\Pi_k^{K}v$
\cite{BeiraoBrezziCangianiManziniEtAl2013,BeiraoBrezziMariniRusso2014,AhmadAlsaediBrezziMariniEtAl2013,AyusodeDiosLipnikovManzini2016}
are used, where $Q_{k-1}^{K}$ is the $L^2$-orthogonal projector onto the
$(k-1)$-th order polynomial space $\mathbb P_{k-1}(K; \mathbb{R}^d)$, and
$\Pi_k^{K}$ is the $H^1$ projection operator onto the $k$-th order polynomial
space $\mathbb P_{k}(K)$.  While only
\[
\|Q_{k-1}^{K}\nabla v\|_{0,K}\lesssim \|\nabla v\|_{0,K}, \quad \|\nabla\Pi_k^{K}v\|_{0,K}\lesssim \|\nabla v\|_{0,K}
\]
hold
rather than the norm equivalence \eqref{intro:gradVknormequiv}, then the additional stabilization term is usually required to ensure the coercivity of the discrete bilinear form.
To remove the additional stabilization term, $\mathbb{V}(K)$ is taken as $\mathbb P_{l}(K; \mathbb{R}^d)$ with large enough $l$ even for the lowest order case $k=1$ in \cite{BerroneBorioMarcon2021,BerroneBorioMarcon2022,DAltriMirandaPatrunoSacco2021}, and the virtual element space $V_k(K)$ has to be modified accordingly to make $Q_{l}^{K}\nabla v$ computable.
Instead, \XH{based on a regular simplicial tessellation of polytope $K$}, we employ $k$-th order or $(k-1)$-th order $H(\div)$-conforming macro finite elements as $\mathbb{V}(K)$ in this paper, and keep the virtual element space $V_k(K)$ as the usual ones.

@@

\subsection{Notation}
Let $\Omega\subset \mathbb{R}^d$ be a bounded
polytope. Given a bounded domain $K\subset\mathbb{R}^{d}$ and a
non-negative integer $m$, let $H^m(K)$ be the usual Sobolev space of functions
on $K$.
The corresponding norm and semi-norm are denoted respectively by
$\Vert\cdot\Vert_{m,K}$ and $|\cdot|_{m,K}$. By convention, let $L^2(K)=H^0(K)$. 
Let $(\cdot, \cdot)_K$ be the standard inner product on $L^2(K)$. If $K$ is $\Omega$, we abbreviate
$\Vert\cdot\Vert_{m,K}$, $|\cdot|_{m,K}$ and $(\cdot, \cdot)_K$ by $\Vert\cdot\Vert_{m}$, $|\cdot|_{m}$ and $(\cdot, \cdot)$,
respectively. Let $H_0^m(K)$ be the closure of $\mathcal C_{0}^{\infty}(K)$ with
respect to the norm $\Vert\cdot\Vert_{m,K}$, and $L_0^2(K)$ consist of all functions in $L^2(K)$ with zero mean value.
For integer $k\geq0$,
notation $\mathbb P_k(K)$ stands for the set of all
polynomials over $K$ with the total degree no more than $k$. Set $\mathbb P_{-1}(K)=\{0\}$.
For a Banach space $B(K)$, let $B(K; \mathbb{X}):=B(K)\otimes\mathbb{X}$ with $\mathbb{X}=\mathbb{R}^d$ and $\mathbb{K}$ being the set of antisymmetric matrices.
Denote by $Q_k^{K}$ the $L^2$-orthogonal projector onto $\mathbb P_k(K)$ or $\mathbb P_{k}(K; \mathbb{X})$.
For tensor $\boldsymbol{\tau}$, let $\skw\boldsymbol{\tau}:=(\boldsymbol{\tau}-\boldsymbol{\tau}^{\intercal})/2$ be the antisymmetric part of $\boldsymbol{\tau}$.
Denote by $\#S$ the number of elements in a finite set $S$.

@@

By (3.5) in \cite{ChenHuang2021divdiv}, we have the direct decomposition
\begin{equation}\label{eq:Pkvectordecomp}  
\mathbb P_{k-1}(T;\mathbb R^{d})=\nabla\mathbb P_k(T)\oplus\mathbb P_{k-2}(T;\mathbb K)\boldsymbol{x},
\end{equation}
\XH{where $\mathbb P_{k-2}(T;\mathbb K)\boldsymbol{x}:=\{\boldsymbol{\tau}\boldsymbol{x}: \boldsymbol{\tau}\in \mathbb P_{k-2}(T;\mathbb K)\}$ with $\boldsymbol{x}\in T$ being the independent variable.}
\XH{Recall that \cite[(35)]{Chen;Huang:2020Finite}
\begin{equation}\label{eq:20230205}
\mathbb P_{k}(T)\cap\ker(I+\boldsymbol{x}\cdot\nabla)=\{0\}.
\end{equation}
% Indeed, \eqref{eq:20230205} follows from the Euler's formula \cite[(2.1)]{ChenHuang2021divdiv}
% $$
% \boldsymbol{x}\cdot\nabla q=kq\quad\forall~q\in \mathbb H_k(T),
% $$
% where $\mathbb H_k(T)$ is the space of homogeneous polynomials of degree $k$.
}

@@

The rectangular domain $\Omega$ is partitioned by the convex polygon mesh
$\mathcal T_0$ and non-convex polygon mesh $\mathcal T_1$, respectively, 
In both virtual element methods \eqref{eq:vem} and \eqref{eq:cfmvem}, we choose $k = 1, 2, 5$.
The numerical results of the nonconforming virtual element method \eqref{eq:vem} on meshes $\mathcal T_0$ and $\mathcal T_1$ are shown in Figure \ref{fig:rate1}. We can see that $\|u - Q_h u_h\|_0=O(h^{k+1})$ and $\|\nabla u - Q_{h, k-1}^{\div}\nabla_h u_h\|_0=O(h^{k})$, which coincide with Theorem~\ref{thm:errorestimateH1}.
And the numerical results of the conforming virtual element method \eqref{eq:cfmvem} are presented in Figure \ref{fig:rate2}. Again $\|u - Q_h u_h\|_0=O(h^{k+1})$ and $\|\nabla u - Q_{h, k}^{\div}\nabla u_h\|_0=O(h^{k})$, which confirm the
theoretical convergence rate in Theorem~\ref{thm:cfmerrorestimateH1}.
\begin{figure}[htbp]
\centering
\begin{minipage}[t]{0.49\linewidth}
\centering
\includegraphics[width=6cm]{./figures/ncvem_convex.pdf}
%\caption{Numberical result of example with $\mathcal T_0$.}
\end{minipage}%
\begin{minipage}[t]{0.49\linewidth}
\centering
\includegraphics[width=6cm]{./figures/ncvem_nonconvex.pdf}
%\caption{[Numberical result of example with $\mathcal T_1$].}
\end{minipage}%
\centering
\caption{Errors $\|u - Q_h u_h\|_0$ and $\|\nabla u - Q_{h, k-1}^{\div}\nabla_h u_h\|_0$ 
of nonconforming virtual element method \eqref{eq:vem} on 
$\mathcal T_0$(left) and $\mathcal T_1$(right) with $k=1, 2, 5$.}
\label{fig:rate1}
\end{figure}
